\chapter{Discussion}
\section{Expressive Motion and Emotion Recognition}
This chapter aims to discuss and interpret the findings presented in the results chapter while also examining them in the context of the research objectives defined in Chapter 1 of the thesis. 
This section will initially examine the emotion recognition results, looking further into which animations succeeded, which fell short, and what this means for the goal of communicating emotion through biped motion. Furthermore, the Likert and Godspeed results are discussed, with an emphasis on how overall perceptions of the robot inform its suitability for human-robot interaction. Where relevant, differences between video-based and live results are examined, and possible explanations are considered.

\section{Emotion recognition}
The four strongest performing animations include:  \textit{Happiness}, 
\textit{Sadness}, \textit{Anger}, and \textit{Fear}. This is likely not a coincidence, as all these emotions can be classified as basic, universal human emotions within Paul Ekman's framework \cite{ekman1999basic}. It is stated further that these basic emotions are associated with distinctive, cross-culturally recognized physical expressions, which may explain why many participants reliably classified them across both study groups. The work of Beck et al. further supports this suggestion \cite{beck_interpretation_2010}, who demonstrated that happiness, sadness, and fear could be successfully communicated through a humanoid robot's body language in the absence of facial expressions, concluding that the absence of facial expression is not a barrier for emotional communication.

However, \textit{Fear} displays a notable exception to this pattern by achieving a recognition rate of 81\%, the 
video-based study scored considerably lower at only 36\%, representing one of the largest gaps between the two groups. This suggests that, although fear has a well-established physical pattern, the animation may rely on subtler motion cues that could be harder to perceive in video. Another reason for this discrepancy may be how the video demonstration was filmed. The videos offer only a fixed perspective for evaluation, but participants in the live trial had the option to move and observe the robot from different angles, potentially providing more context. This is also in line with the work of \cite{beck_interpretation_2010}, which notes that fear was among the harder emotions to recognize, suggesting that, despite being a basic emotion, its physical expression may be more context-dependent and sensitive to observational conditions.

The other emotion/intent animations, including \textit{Curiosity} and \textit{Frustrated}, are not considered basic emotions in Ekman's framework. They are more complex, socially constructed states, possibly lacking the well-known hardwired physical expressions associated with basic emotions. This could explain why these animations were perceived as more ambiguous across both study groups. 

\textit{Curiosity} was the weakest performing animation overall, achieving consistently low recognition rates in both study groups. Unlike happiness or anger, curiosity is more of a cognitive state and does not offer a strongly defined set of physical expressions. This animation was designed to simulate an intention rather than an emotional state. making it more difficult to communicate through body language alone. The high concentration of \textit{other} responses suggests that participants perceived the motion without being able to map it to a specific emotion or intent. This is also consistent with the findings in \cite{beck_interpretation_2010}, where recognition rates dropped noticeably for emotional states lacking clear physical expression patterns. Improving the curiosity animation may require additional expressive actions, such as clearer use of the eye displays and antenna movement. 

The \textit{Frustrated} animation presented a different challenge as it had a larger concentration of \textit{adjacent} responses rather than being misclassified entirely. Consistent adjacent responses include disgust, impatience, and being upset. This suggests that the animation successfully communicated a negative state even if the intended emotion was not identified. Frustration also shares physical characteristics with anger and disgust, both of which are basic emotions in Ekman's framework \cite{ekman_basic_1999}, which may explain why participants reliably perceived it as a negative state. The large gap between recognition rates in live and video-based trials further suggests that the frustrated animation relies on subtler physical cues that are more easily picked up in a live demonstration.

\section{Live vs Video Study Differences}
The overall trend in the studies' results is that the live-based approach consistently outperformed the video-based approach in recognition rates and Likert-scale results. Based on the study's design, several factors may explain the performance gaps between the two study groups.
The large performance gaps may be a result of the observation methods applied. Observing the robot live, being able to witness the physical attributes like sound, size, and mechanics, could provide participants with additional information, making it easier to interpret the intended emotions or intentions of the robot platform. The video-based participants were presented with a fixed, top-down view of the robot, which may have limited their field of view, depth perception, and ability to observe subtle cues, making the animations less clear. With this in mind, it could be argued that observing the robot physically granted participants an advantage in emotion recognition, causing them to give higher, more accurate scores. 
Another possible reason for the variations is the large differences in the demographics of the two studies. The live study had a participant composition in which 88\% were engineers, compared to 50\% in the video-based study. The Live study had a larger population of participants reporting a high level of robot familiarity. Given that the live study consisted mostly of engineers with greater familiarity with robots, it could be argued that the participants would be more accustomed to robot motions, leading them to rate the robot higher.
The reason for the live-based study consistently outperforming the video-based study is likely a combination of the 2 factors discussed above. The fixed camera perspective may have obscured subtle motion cues that would have been more apparent in person. At the same time, the higher proportion of engineering students in the live group may have made them more attuned to interpreting the mechanical motion.

\section{Overall Robot Perception}
The godspeed questionnaire approach provided a broader view of overall perceptions of the robot beyond individual animations. While all five animations offer valuable insight, three of them are particularly significant for the goal of this thesis:  \textit{likability}, \textit{perceived safety}, and \textit{animacy}. The robot platform was designed with human-robot interactions in mind; the robot needed to be perceived as likable and non-threatening. If these dimensions all scored poorly, it would fail in its core purpose regardless of how well the individual animations performed. Animacy was also considered more relevant as it directly reflects whether the expressive motion framework produced a robot that comes across as lifelike.

\textit{Likability} was the highest scoring dimension, with mean scores of 4.50 and 4.45 in the live and video-based studies, respectively. This is especially noteworthy, as it was one of the most important dimensions. Having a well-liked robot can only enhance the potential for pleasant robot-human interactions. This could also suggest that the expressive motion framework was successful in creating a likable and engaging robot platform, independent of whether participants correctly classified the intended emotions.

\textit{Perceived Safety} also scored high in both groups, with mean scores of 4.26 and 4.19 in live and video-based trials. This suggests that participants felt safe and comfortable in the robot's presence. This is a meaningful result for a platform intended for close human interaction; the robot platform's small, non-threatening form factor appears to have contributed positively to this perception.

\textit{Animacy} scored around the midpoint in both groups (3.40 live, 3.49 video), suggesting that participants perceived the robot as moderately lifelike.  This is not a particularly high score, but it indicates that the expressive framework succeeded in creating some sense of life in the robot, even in the absence of complex facial expressions or humanoid features. This also highlights a potential weakness in the expressive motion framework, as this score should have been higher in a robot designed for expressive motion in human-robot interactions. The lower score may be due to the movements being perceived as more mechanical than organic, with abrupt transitions between animations.

\textit{Anthropomorphism} received the lowest scores across both groups (3.14 live, 3.15 video), which is expected, as it is consistent with the design choices that deliberately avoid humanoid features. Having a robot platform that scores low on anthropomorphism but high on likability suggests that human-like appearance is not a prerequisite for positive perception or good human-robot interactions. It can also be argued that expressive motion alone can be sufficient to create an engaging and likable robot platform.

\textit{Perceived Intelligence} is among the lower-scoring dimensions (3.20 live, 3.45 video), which is consistent with the design goals of the platform. Miwo was never intended to perform complex or logical tasks, but rather to function as an expressive companion. Scores around the midpoint suggest that participants perceived the robot as moderately intelligent, which is a reasonable and expected result for a platform of this kind.

A notable trend across all five dimensions is the consistency between the live and video-based studies. This is interesting, as the two groups showed large gaps in their responses in the study's emotion recognition section. This suggests that the observation method significantly affects participants' ability to classify emotions and intentions, but has little to no impact on the overall perception of the robot platform.

\section{Limitations}
While the results of the human perception study provide valuable insight into the expressive capabilities of the Miwo platform, several limitations should be acknowledged when interpreting the findings.
Significant limitations include the relatively small sample size: 38 participants (16 live, 22 video) completed either the live or video-based study. The results show clear differences and trends, but a larger, more diverse participant population would be needed to draw stronger conclusions.

Other limitations include the demographic bias present in both groups, particularly in the live group. Participants with engineering backgrounds were overrepresented in both studies, comprising 88\% in the live study and 50\% in the video-based study. This bias could have influenced the recognition rates, as participants in engineering may be more accustomed to interpreting robotic motion. The results may therefore not be fully representative of how the general population would perceive the robots' expressive capabilities.

The order of animations in both trials was also fixed and identical for all participants, which could introduce ordering effects. Participants may have become more familiar with the robot's movements as the study progressed, potentially influencing their evaluation of later animations. Participants could also develop fatigue, causing them to evaluate later animations differently from those initially presented.

Finally, the video-based study used recordings from a fixed camera perspective, limiting participants to a single viewpoint. As discussed in the previous section, this may have made more subtle motion cues less apparent than they would have been in person. This could also have contributed to the overall lower recognition rates observed in the video-based study.

The final animations were not produced by professionals but by a student with no prior knowledge or experience in animation. Therefore, the animations are rather simple, and more time and care could have produced better results, more in line with professional animation.

